On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-10 Revisiones/PNA_MMP_Responses to the reviewers.tex

%\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
%\tableofcontents

We  are grateful to the reviewers for taking their time  to review our work. Their comments have helped us to improve the paper. Changes to the original manuscript are given in  color \textcolor{red}{red} for corrections and  \textcolor{blue}{blue} for modified text in the revised version.  A detailed response to both reviewers is given below.

@@

\textit{Technical issues for the authors follow:}
 \begin{enumerate}
\item \textit{Cell-crossing is a well known issue with the MPM but in the review no mention is made of CPDI approaches.}\\

A citation of the work of \cite{Sadeghirad_2011} with the CPDI has been added to the state of art.

\item \textit{The development is very confusing and would make more sense were it to be tied to 
the way that GIMP methods are usually stated, i.e. that the standard FE shape functions are replaced by weighting functions comprised of the conjugation of a standard shape function and a particle "characteristic" function. Line 155 demonstrates this, i.e. the use of $N_{Ip}$ before explanation.}\\

Unfortunately, the \textit{max-ent} approximation can not be expressed as the conjugation of a standard shape function and a particle "characteristic" function. However, in order to improve the readability of the MPM development, the approach has been revised with a detailed explanation.

@@

\item \textit{There is little discussion on some of the pitfalls of using max-ent, e.g. the need for a convex domain, the issue of determining Lagrange Multiplier solutions for points close to boundaries, i.e. the solutions process here at Eqns 37 and 38.}\\

Of course, the employment of the \textit{max-ent} approximants is not free of difficulties. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of it a challenging tool. However, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical MPM problems such as cell-crossing, stress instabilities and the use of unstructured mesh.

Concerning the first issue mentioned by the reviewer, the special case of non-convex domains was discussed by Arroyo \& Ortiz~\cite{Arroyo2006}. Some of the solutions proposed by the aforementioned authors are : the possibility of replacing the Euclidean distance $\lVert  x - x_a  \rVert$ in the definition of the shape functions by the length of the shortest path contained within the domain connecting $x$ and $x_a$. Or the decomposition of the non-convex domain into convex sub-domains.  This topic has also been extensively studied in the context of MLS-based meshfree methods, for instance visibility, diffraction, and constrained path criteria. These methods are directly applicable to local \textit{max-ent} approximation.

@@

\item \textit{Does the use of max-ent solve the problem of imposing Dirichlet boundary conditions in the MPM?}\\

Unfortunately, local \textit{max-ent} approximation technique only provides an efficient tool to transfer information from particles to nodes and vice versa to the MPM. Hence, the difficulties of imposing Dirichlet boundary conditions (DBC) in the MPM persists since they are a direct consequence of discretising the continuum with material points instead of nodes, where traditionally DBC are imposed. However, any available technique employed to overcome this situation in the standard MPM can be translated to the \textit{max-ent} one thanks to the Kronecker-Delta property inherent to it \cite{Arroyo2006}. See for instance the work of Cortis {\it et al.} (2018) \cite{Cortis_et_al_2017_IJNME} as example of how DBC can be imposed in the MPM.

@@

\item \textit{I am sceptical of the novelty of the proposed explicit scheme as there is no mention of the recent XPIC(m) approach of Hammerquist and Nairn or similar approaches used in the Computational Graphics community, e.g. Stomakhin et al. (2013) or the more recent Polypic.}\\

These interesting publications will be added to the state of art of the revised manuscript.In the humble opinion of the authors, the XPIC approach is dedicated to mitigate the null space error produced during the transfer of information between particles and nodes. Therefore, it is related to the spatial discretisation in stead of temporal. Meanwhile, the algorithm introduced in this research proposes the stabilisation within the adaptation of the predictor-corrector algorithm, which has been used traditionally in FEM to deal with fast dynamics. To the author\textquotesingle s knowledge, this is the first attempt to adapt the  traditional predictor-corrector for the MPM framework.

@@

\end{enumerate}

 \textit{The numerical examples are not impressive when the message of the paper is that this is good for fast dynamics problems. Compare to papers such as Ma et al. (2009) IJ Impact Engineering}\\

Contrary to the reviewer\textquotesingle s opinion, the provided results are capable to highlight the strong capabilities of the MPM Newmark Predictor-Corrector scheme (NPC). It is important to remark the extension of this research: in order to isolate the strength of the proposed methodology, covering spatial and time discretization, only linear elastic problems are assessed.

Although the results presented in Ma {\it et al.} \cite{MA2009272} are impressive, the proposed examples are out of the scope of the present research. In order to reinforce the benefits of the proposed methodology, we propose a 2D wave propagation problem has been simulated. It is analogous to the work of Hammerquist \& Nairn \cite{HAMMERQUIST2017724} (previously cited by the reviewer). This new test highlights the capabilities of the NPC to reproduce complex wave patterns under challenging loads such as impacts. Which is more, it is able to reduce the presence of numerical noise compared with the FE solution. Additionally, the robustness of the local \textit{max-ent} approximation is exposed dealing with unstructured mesh.

### 2020-09-10 LME-MPM.tex
Change in figure 14
Cambios en la hoja de respuestas

\textcolor{blue}{The following test is intended to validate the ability of the discussed time integration schemes in order to reproduce the wave propagation in 2D. The impact of an elastic square plate against a vertical wall is assessed. The square plate dimensions are 50 x 50 m and their elastic parameters are:
\begin{itemize} 
\item  Initial density : $20\ kg/m^3$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $0.1\ MPa$
\end{itemize}
}
\textcolor{blue}{The background set of nodes is composed by a cartesian grid with an uniform spatial discretisation of 1 m. Four material points per cell are considered. The Courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{3.0}$ shape functions are employed. A horizontal velocity of the square plate of $\text{V}_0 = 1.0 m/s$ until the shock against the obstacle, which has been modelled as a set of nodes with prescribed velocity equal to zero. Stress waves are produced by the aforementioned shock what causes the square to bounce back. Both \acrshort{NPC} and \acrshort{FE} schemes are compared. Spatial distribution of the velocity field inside the square plate obtained with both schemes at different times are given in figure ~\ref{fig:Magnitude_velocity_impact_square}}. 
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=1]{Figures/Velocity_FE_vs_NPC_square.png} 
  }
  \caption{Spatial distribution of the magnitude of the velocity field in an elastic square plate impacted on the right edge. The top row depicts the \acrshort{npc} results meanwhile the bottom row represents the \acrshort{fe} ones at different instants.}
  \label{fig:Magnitude_velocity_impact_square}
\end{figure}
%%%%%%%%%%%%%%%% 
\textcolor{blue}{At early times, \acrshort{npc} and \acrshort{fe} behvae in a similar manner. However, at later times, the \acrshort{fe} scheme introduces significant numerical noise compared with the \acrshort{npc} one. At the end of the simulation, the filtering process when \acrshort{npc} is used leads to a negligible loss of mechanical energy (about 0.1\%).}

\textcolor{blue}{Finally, an unstructured background mesh is proposed in order to validate the performance of the \acrshort{lme} shape function against the \acrshort{ugimp} one.}
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=0.8]{Figures/Velocity-structured-unstructured-lme-ugimp.png}
  }
  \caption{Sensitivity analysis of \acrshort{ugimp} and \acrshort{lme} schemes to the randomness in the layout of the background set of nodes. On the left column, sketches of the two employed meshes are depicted. On the right side, spatial distribution of the velocity magnitudes are shown for each case. All the simulation where carried out with the proposed \acrshort{npc} scheme.}
  \label{fig:Magnitude_velocity_impact_square_structured_unestructured}
\end{figure}
%%%%%%%%%%%%%%%%
\textcolor{blue}{ As it is observed in figure~\ref{fig:Magnitude_velocity_impact_square_structured_unestructured}, the \acrshort{lme} approximation preserves the solution even though the mesh is not structured. Meanwhile, the \acrshort{ugimp} is not able to deal with the unstructured mesh as expected. %It goes without saying that there is no point in using unstructured meshes in this type of situation. 
%Despite this, 
Thus, the \acrshort{lme} approach shows remarkable robustness in dealing with an unstructured mesh. This is useful when more complex domains are to be discretised since the employment of  structured mesh is not possible.}

>>>>

\textcolor{blue}{The following test is intended to validate the ability of the discussed time integration schemes in order to reproduce the wave propagation in 2D. The impact of an elastic square plate against a vertical wall is assessed. The square plate dimensions are 50 x 50 m and their elastic parameters are:
\begin{itemize} 
\item  Initial density : $20\ kg/m^3$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $0.1\ MPa$
\end{itemize}
}
\textcolor{blue}{The background set of nodes is composed by a cartesian grid with an uniform spatial discretisation of 1 m. Four material points per cell are considered. The Courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{3.0}$ shape functions are employed. A horizontal velocity of the square plate of $\text{V}_0 = 1.0 m/s$ until the shock against the obstacle, which has been modelled as a set of nodes with prescribed velocity equal to zero. Stress waves are produced by the aforementioned shock what causes the square to bounce back. Both \acrshort{npc} and \acrshort{fe} schemes are compared. Spatial distribution of the velocity field inside the square plate obtained with both schemes at different times are given in figure ~\ref{fig:Magnitude_velocity_impact_square}}. 
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=1]{Figures/Velocity_FE_vs_NPC_square.png} 
  }
  \caption{Spatial distribution of the magnitude of the velocity field in an elastic square plate impacted on the right edge. The top row depicts the \acrshort{npc} results meanwhile the bottom row represents the \acrshort{fe} ones at different instants.}
  \label{fig:Magnitude_velocity_impact_square}
\end{figure}
%%%%%%%%%%%%%%%% 
\textcolor{blue}{At early times, \acrshort{npc} and \acrshort{fe} behvae in a similar manner. However, at later times, the \acrshort{fe} scheme introduces significant numerical noise compared with the \acrshort{npc} one. At the end of the simulation, the filtering process when \acrshort{npc} is used leads to a negligible loss of mechanical energy (about 0.1\%).}

\textcolor{blue}{Finally, an unstructured background mesh is proposed in order to validate the performance of the \acrshort{lme} shape function against the \acrshort{ugimp} one.}
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=0.8]{Figures/Velocity-structured-unstructured-lme-ugimp.png}
  }
  \caption{Sensitivity analysis of \acrshort{ugimp} and \acrshort{lme} schemes to the randomness in the layout of the background set of nodes. On the left column, sketches of the two employed meshes are depicted. On the right side, spatial distribution of the velocity magnitudes are shown for each case at second 2.66. All the simulation where carried out with the proposed \acrshort{npc} scheme.}
  \label{fig:Magnitude_velocity_impact_square_structured_unestructured}
\end{figure}
%%%%%%%%%%%%%%%%
\textcolor{blue}{ As it is observed in figure~\ref{fig:Magnitude_velocity_impact_square_structured_unestructured}, the \acrshort{lme} approximation preserves the solution even though the mesh is not structured. Meanwhile, the \acrshort{ugimp} is not able to deal with the unstructured mesh as expected. %It goes without saying that there is no point in using unstructured meshes in this type of situation. 
%Despite this, 
Thus, the \acrshort{lme} approach shows remarkable robustness in dealing with an unstructured mesh. This is useful when more complex domains are to be discretised since the employment of  structured mesh is not possible.}
